---
title: 'MALDI-informed microdissected Airway proteomics'
author: Geremy Clair
date: October 19, 2022
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    theme : cerulean
    code_folding: hide
---

```{r knit_options, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
knitr::opts_chunk$set(fig.width=10, fig.height=6, echo = TRUE, results = TRUE, warning = FALSE, message=FALSE) 
options(repos=structure(c(CRAN="https://mirror.ibcp.fr/pub/CRAN/")))
setwd("~/Documents/GitHub/MALDI_informed_microdissected_Airway_proteomics/")
```

# Required libraries

The required libraries are loaded 
- RomicsProcessor written by Geremy Clair (2022) is used to perform trackable transformation and statistics to the dataset 
```{r load_packages}
library("RomicsProcessor")
```


# MaxQuant import

The iBAQ data contained in the protein table was loaded, the corresponding metadata was loaded
```{r data_metadata_load}
data<-extractMaxQuant("./01_source_files/proteinGroups.txt",quantification_type = "iBAQ",cont.rm = T,site.rm = T,rev.rm = T)
IDsdetails<-extractMaxQuantIDs("./01_source_files/proteinGroups.txt",cont.rm = T,site.rm = T,rev.rm = T)
IDsdetails<-cbind(UniProt_Name=sub(".*\\|","",IDsdetails$protein.ids), IDsdetails)
metadata<- read.csv(file = "./01_source_files/metadata.csv")
colnames(data)<- sub("iBAQ.","",colnames(data))
colnames(metadata)<-tolower(colnames(metadata))
write.csv(IDsdetails,"03_output_files/MaxQuantIDS.csv")
```

# Romics_object creation

The data and metadata were placed in an romics_object, the sample names were retrieved from the metadata
```{r create_romics}
romics_proteins<- romicsCreateObject(data, metadata,main_factor = "desc")
```

# Data evaluation and transformation

The missingness was evaluated for each section
```{r missingness_eval}
romics_proteins<- romicsZeroToMissing(romics_proteins)
romicsPlotMissing(romics_proteins)
```

The proteins to be conserved for quantification were selected to contain at least 70% of complete values (5/7 samples), the overall missingness was evaluated after filtering.
```{r missing_filter}
romics_proteins<-romicsFilterMissing(romics_proteins, percentage_completeness = 70,main_factor = "none")
print(paste0(nrow(romics_proteins$data),"/", nrow(romics_proteins$original_data)," proteins remained after filtering", " (",round(nrow(romics_proteins$data)/nrow(romics_proteins$original_data)*100,2),"%)."))
romicsPlotMissing(romics_proteins)
```
The protein abundance was median centered prior to plot distribution boxplots.
```{r median_norm}
romics_proteins<-log2transform(romics_proteins)
romics_proteins<-medianCenterSample(romics_proteins)
distribBoxplot(romics_proteins)
```

The grouping of the samples by is checked by hierarchical clustering and PCA.
```{r hclust_no_imput}
romicsHclust(romics_proteins)
indPCAplot(romics_proteins,Xcomp = 1,Ycomp = 2,plotType = "individual",ncp=5)
```

Finally the data was exported.
```{r export_results}
results<-romicsExportData(romics_proteins,statistics = F,missing_data = F)
results<-merge(IDsdetails,results,by.x = "UniProt_Name", by.y=0)
write.csv(results, "./03_output_files/results_MicroPOTS_proteomics.csv")
```
